### 2020-08-11 r-package/tests_marcus/rjava_tests.Rmd
Added loadDestinationPointsFromCsv method to r5r api, and succesfully called it from R

---
title: "Package Tests"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library("rJava")
```

## Testing Java / R Integration

```{r}
.jinit()
.jaddClassPath(path = "../../JAR/R5.jar")
.jaddClassPath(path = "../../JAR/r5r_core.jar")
.jclassPath()
```

```{r}
r5r_core <- .jnew("R5RCore", here::here("inst/extdata/"))
```

```{r}
.jcall(r5r_core, returnSig = "V", 
       method = "loadDestinationPointsFromCsv", here::here("inst/extdata/", "poa_hexgrid.csv"))
```