### 2020-08-13 r-package/tests_marcus/rjava_tests.Rmd
Added single trip planning to r5r_core Tested single trip planning from R using .jcall

---
title: "Package Tests"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library("rJava")
library("jdx")
library("sf")
library("dplyr")
library("ggplot2")
```

```{r}
poa_boundary <- st_read(here::here("inst/extdata/portoalegre_boundary.geojson"))
```

## Testing Java / R Integration

```{r}
.jinit()
.jaddClassPath(path = "../../JAR/R5.jar")
.jaddClassPath(path = "../../java-api/out/artifacts/r5r_core/r5r_core.jar")
.jclassPath()
```

```{r}
r5r_core <- .jnew("R5RCore", here::here("inst/extdata/"))
```

```{r}
.jcall(r5r_core, returnSig = "V", 
       method = "loadDestinationPointsFromCsv", here::here("inst/extdata/", "poa_hexgrid.csv"))
```

Estação Farrapos
Lat: -29.997611
Lon: -51.197720

Shopping Praia de Belas
Lat: -30.048951
Lon: -51.229533

Function Signature
public void planSingleTrip(double fromLat, double fromLon, double toLat, double toLon,
                               String[] directModes, String[] transitModes, String date, String departureTime) {

```{r}
fromLat <- -29.997611
fromLon <- -51.197720
toLat <- -30.048951
toLon <- -51.229533
trip_date <- "2019-04-08"
departure_time <- "14:00:00"

.jcall(r5r_core, returnSig = "V", 
       method = "planSingleTrip", fromLat, fromLon, toLat, toLon, "WALK", "BUS", trip_date, departure_time)
```

```{r}
path_options <- r5r_core$getPathOptionsTable()
path_options_df <-convertToR(path_options)
```

```{r}
path_options_df %>% 
  filter(option <= 4) %>%
  mutate(geometry = st_as_sfc(geometry)) %>%
  st_sf(crs = 4326) %>%
  ggplot() +
  geom_sf(data=poa_boundary) +
  geom_sf(aes(geometry=geometry, colour = mode)) +
  scale_colour_brewer(palette = "Set1") +
  facet_wrap(~option, ncol = 4) +
  theme(legend.position = "bottom")
```