### 2020-08-24 r-package/doc/intro_to_r5r.Rmd

---
title: 'r5r: Rapid Realistic Routing with R5 in R'
author: "Rafael H. M. Pereira, Marcus Saraiva, Daniel Herszenhut, Carlos Kaue Braga"
date: "`r Sys.Date()`"
output:
  rmarkdown::html_vignette: default
  github_document: default
abstract: "`r5r` is an R package for rapid realistic routing on multimodal transport networks (walk, bike, public transport and car) using R5. The package allows users to generate detailed routing analysis or calculate travel time matrices using seamless parallel computing on top of the R5 Java machine <https://github.com/conveyal/r5>"
urlcolor: blue
vignette: >
  %\VignetteIndexEntry{r5r-intro} 
  %\VignetteEncoding{UTF-8}
  %\VignetteEngine{knitr::rmarkdown} 
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

# Introduction

**r5r** is an [R package for rapid realistic routing on multimodal transport networks](https://github.com/ipeaGIT/r5r) (walk, bike, public transport and car).
It provides a simple and friendly interface to R5, a really fast and open source 
routing engine based in `Java` developed by [Conveyal](https://www.conveyal.com/).
R5 stands for [Rapid Realistic Routing on Real-world and Reimagined networks](https://github.com/conveyal/r5).

# Installation
To use `r5r`, you need to have [Java SE Development Kit 11.0.8](https://www.oracle.com/java/technologies/javase-jdk11-downloads.html) 
installed on your computer. No worries, it's free.

Install the development version of `r5r` from GitHub (soon on CRAN) and load packages.

```{r, eval = FALSE}
devtools::install_github("ipeaGIT/r5r", subdir = "r-package")

```

```{r, message = FALSE}
library(r5r)
library(sf)
library(data.table)
library(ggplot2)
library(mapview)

```

The package has 3 fundamental functions.

1. `setup_r5` to build a routable transport network given a street and public
transport networks

2. `travel_time_matrix` for fast computation of travel time estimates between 
origin destination pairs

3. `detailed_itineraries` to get detailed information on one or multiple 
alternative routes between origin destination pairs

Let's have a quick look at how these functions work using a sample data set.

# Demonstration on sample data

## Data
To illustrate functionality, the package includes a small sample data for the 
city of Porto Alegre (Brazil). It includes three files:

* An Open Street Map network in `.pbf` format (*mandatory*)
* A public transport network in `GTFS.zip` format
* The spatial coordinates of the origin destination pairs in a `.csv` file.

```{r}
data_path <- system.file("extdata", package = "r5r")
list.files(data_path)

```
The data with origin destination pairs looks like this below. In this example, 
we will be looking at the transport routes between ten random points in this 
data set.
```{r}
points <- fread(system.file("extdata/poa_hexgrid.csv", package = "r5r"))
points <- points[ c(sample(1:nrow(points), 10, replace=TRUE)), ]
head(points)

```
#### Set memory

Before we start, we need to increase the memory available to Java. This is
because by default `R` allocates only 512MB of memory for Java processes, which is
would not b enough for large queries using `r5r`. To increase available memory 
to 2GB, for example, we need to set the java.parameters option at the beginning 
of the script, as follows:

```{r, message = FALSE, eval = FALSE}
options(java.parameters = "-Xmx2G")
```

## Build routable transport network with `setup_r5()`

The first step is to build the multimodal transport network used for routing in 
R5. This is done with the `setup_r5` function. This function does two things: (1)
downloads/updates a compiled JAR file of R5 and stores it locally in the `r5r` package 
directory for future use; and (2) combines the osm.pbf and gtfs.zip data sets to
build a routable network object.

```{r, message = FALSE, eval = FALSE}
# Indicate the path where OSM and GTFS data are stored
r5r_core <- setup_r5(data_path = data_path, verbose = FALSE)

```

## Routing analysis

For fast routing analysis, **r5r** currently has two core functions: `travel_time_matrix` 
and `detailed_itineraries`.

### Fast many to many travel time matrix

The `travel_time_matrix` function is a really simple and fast function to compute
travel time estimates between one or multiple origin destination pairs. The 
origin/destination input can be either a spatial `sf POINT` object, or a 
`data.frame` containing the columns `id, lon, lat`.

```{r, message = FALSE, eval = FALSE}
# calculate a travel time matrix
ttm <- travel_time_matrix( r5r_core = r5r_core,
                           origins = points,
                           destinations = points,
                           departure_datetime = lubridate::as_datetime("2019-03-20 14:00:00"),
                           mode = c("WALK", "TRANSIT"),
                           max_walk_dist = 5,
                           max_trip_duration = 7200,
                           verbose = FALSE
                          )
head(ttm)
```

```{r ttm head, echo = FALSE, message = FALSE}
knitr::include_graphics(system.file("img", "vig_output_ttm.png", package="r5r"))
```

### Detailed itineraries

Most routing packages only return the fastest route. A key advantage of the 
`detailed_itineraries` function is that is allows for fast routing analysis while
providing multiple alternative routes between origin destination pairs. The output
also brings detailed information for each route alternative at the trip segment
level, including the transport mode, waiting times, travel time and distance of
each trip segment.

In this example below, we want to know some alternative routes between one origin
destination pair only.

```{r, message = FALSE, eval = FALSE}
# inputs
points <- read.csv(file.path(data_path, "poa_points_of_interest.csv"))
origins <- points[10,]
destinations <- points[12,]
mode = c("WALK", "TRANSIT")
max_walk_dist <- 10000
departure_datetime <- lubridate::as_datetime("2019-03-20 14:00:00")

df <- detailed_itineraries(r5r_core = r5r_core,
                         origins,
                         destinations,
                         mode,
                         departure_datetime,
                         max_walk_dist,
                         shortest_path = FALSE,
                         verbose = FALSE)

head(df)
```
```{r detailed head, echo = FALSE, message = FALSE}
knitr::include_graphics(system.file("img", "vig_output_detailed.png", package="r5r"))
```

The output is a `data.frame sf` object, so we can easily visualize the results.

#### Visualize results

**Static visualization** with `ggplot2` package: To facilitate the 
visualization of the results in `ggplot2`, you can also use the 
`street_network_to_sf` function to  extract the OSM street network used in the
routing.

```{r, message = FALSE, eval = FALSE}
# extract OSM network
street_net <- street_network_to_sf(r5r_core)

# plot
ggplot() +
  geom_sf(data = street_net$edges, color='gray85') +
  geom_sf(data = df, aes(color=mode)) +
  facet_wrap(.~option) + 
  theme_void()

```

```{r ggplot2 output, echo = FALSE, message = FALSE}
knitr::include_graphics(system.file("img", "vig_detailed_ggplot.png", package="r5r"))
```

**Interactive visualization** with `mapview`:

```{r, message = FALSE, eval = FALSE}
mapviewOptions(platform = 'leafgl')
mapview(df, zcol = 'option')

```

```{r mapview output, echo = FALSE, message = FALSE}
knitr::include_graphics(system.file("img", "vig_detailed_mapview.png", package="r5r"))
```

If you have any suggestions or want to report an error, please visit the [the package GitHub page](https://github.com/ipeaGIT/r5r).